---
title: "The Maps of New Zealand"
author: "Alfiya Shafy"
date: "2024-09-10"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The Basic Map of New Zealand

Here we create a basic map of New Zealand which can be zoomed in and out to display the small details of the area.

```{r}
library(leaflet)
library(dplyr)
nz_map <- leaflet() %>%
  addTiles() %>%
  setView(lng = 174.885971, lat = -40.900557, zoom = 5)

nz_map
```

## Adding Markers to the Map

In the above map we created we are gonna add markers to mark the major cities in New Zealand such as Auckland, Wellington and Christchruch.

```{r}
#Create New Zealand dataset.
nz_pop <- data.frame(
  city = c("Auckland", "Wellington", "Christchurch"),
  lat = c(-36.85102, -41.29352, -43.53305),
  lng = c(174.75796, 174.77760, 172.63093),
  population = c(1693000, 424000, 408000)
)

#Create New Zealand map with makers.
nz_map_pop <- leaflet(nz_pop) %>%
  addTiles() %>%
  setView(lng = 174.7633, lat = -40.900557, zoom = 5) %>%
  addMarkers(~lng, ~lat, popup = ~paste(city, "<br>Population:", population))

nz_map_pop
```

## Creating the Leaflet Map and Add a Provider Tile

```{r}
# Create the leaflet map
nz_map <- leaflet(nz_pop) %>%
  addTiles() %>%  # Basic tiles to render the map
  setView(lng = 174.7633, lat = -40.900557, zoom = 5) %>%  # Center on New Zealand
  addMarkers(~lng, ~lat, popup = ~paste(city, "<br>Population:", population)) %>%  # Add city markers
  addProviderTiles(providers$CartoDB.Positron)  # Use a clear and readable provider tile

# Display the map
nz_map
```

# Creating Heatmap

Now we create heatmap depicting the population density of major cities in New Zealand.

```{r}
library(leaflet)
library(leaflet.extras)

nz_map_heat <- leaflet(nz_pop) %>%
  addTiles() %>%
  setView(lng = 174.7633, lat = -40.900557, zoom = 5) %>%
  addHeatmap(lng = ~lng, lat = ~lat, intensity = ~population, blur = 20, max = 1, radius = 15)

nz_map_heat

```

# Reading GeoJSON File.
```{r}
library(sf)
nz_geo <- st_read("C:/Users/dell/Downloads/nz_ta.geojson")
```

## Creating Black and White Plot

Now we create a black and white map of the New Zealand territories.

```{r}
library(ggplot2)
ggplot(nz_geo) +
  geom_sf() +
  theme_void()

```

## Loading Population dataset

We now load the population csv file dataset.

```{r}
#Loading the data.
population_data <- read.csv("C:/Users/dell/Downloads/nz_territory_2016_population.csv")

#Removing the unnecessary data points.
names(population_data) <- tolower(names(population_data))
str(population_data)


```

## Merging the Geospatial and Population Data.

```{r}
library(sf)
library(dplyr)

# Perform a left join using dplyr
merged_data <- nz_geo %>%
  left_join(population_data, by = c("TA2016_NAM" = "nz_territory"))

# Check the result
print(merged_data)

#Renaming Colunm names 
library(dplyr)

# Rename columns in merged_data
merged_data <- merged_data %>%
  rename(
    NZ_territory = TA2016_NAM,
    Population = x2016_population
  )
```

# Creating the Basic Choropleth Map

Here we create the basic Choropleth Map.

```{r}
ggplot(merged_data) +
  geom_sf(aes(fill = Population)) +
  theme_void()
```

# Advanced Choropleth Map

Here we create more sophisticated Choropleth Map.

```{r}
nz_plot <- ggplot(merged_data) +
  geom_sf(aes(fill = log(Population))) +
  scale_fill_viridis_c() +
  theme_void() +
  labs(
    title = "New Zealand Population by Territory",
    fill = "Population (Log Scale)"
  )

ggsave("nz_plot.png")

```

# Extra for Experts

We are going to create a more interactive Choropleth Map using RColorBrewer and ploty packages.

```{r}
#Loading packages
library(plotly)
library(RColorBrewer)
library(sf)
library(dplyr)

#Defining Color palette
color_palette <- brewer.pal(9, "Blues")

merged_data$Population <- as.numeric(merged_data$Population)

# Create the interactive map
plot_ly(data = merged_data, 
        type = 'scattermapbox',
        mode = 'markers',
        color = ~Population,
        colors = color_palette,
        text = ~paste("Territory:", merged_data$NZ_territory, "<br>Population:", Population),
        hoverinfo = 'text',
        marker = list(size = 10, opacity = 1)) %>%
  layout(mapbox = list(style = 'open-street-map',
                       center = list(lon = 174.8860, lat = -40.9006),
                       zoom =3
                       ),
         title = 'Interactive Choropleth Map of New Zealand Population',
         coloraxis = list(colorbar = list(title = 'Population'))) %>%
  config(displayModeBar = TRUE)
```

